### 2014-06-20 kypchak-morphfest-2014/poster.tex
a complete draft
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@54688 72bbbca6-d526-0410-a7d9-f06f51895060

	\headerbox{Development effort}{name=devel,column=4,row=0}{
	\scalebox{0.92}{
		\begin{tabular}{lllll}
			\toprule
			 & \textbf{Kyrgyz} & \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
			\midrule
			begun & Apr.\ 2011 & Dec.\ 2010 & Dec.\ 2011 & Oct.\ 2013 \\
			80\% cov.\ & Aug.?\ 2011 & Aug.\ 2012 & Aug.\ 2012 & Oct.\ 2013 \\\midrule
			time & 4 months & 19 months & 7 months & 1 week \\
			\bottomrule
		\end{tabular}
	}
		\begin{itemize}
			%\item Kyrgyz and Kazakh transducers begun in 2011
			%\item Kyrgyz transducer reached 80\% coverage summer of 2011
			\item Kazakh transducer based on Kyrgyz transducer
			\item Kumyk transducer based on Kazakh, Tatar transducers
			\item \tilda{}1 week to reach 80\% coverage, +1 week to reach 90\%
		\end{itemize}
		\vspace*{-2.5em}
		\hspace*{-2em}\includegraphics[width=0.9\textwidth,angle=270]{../kipchak-lrec-2013/graphs/kumyk-yoldash-cov}
		\vspace*{-2em}
	}

@@

		\headerbox{Categorisation}{name=adjectives,column=4,below=devel}{
		%\headerbox{Morphophonology}{name=morphophon,below=morphtrans}{
			\htwo{Adjectives}
			\begin{itemize}
				\item morphologically distinct adjective classes
				\item most sources:\ adjectives can be used substantively and adverbially
				\item Other Turkic transducers:\ \texttt{0}-derivation (\hilitetwo{overgenerates})
				\item but not all adjectives have all of the following:\\
				comparative forms, substantive readings, adverbial readings
				\item Our approach:\ categorisation 
				\item only correct forms are analysed and generated
			\end{itemize}

		\scalebox{0.7}{
			\begin{tabular}{lllll}
				\toprule
				\textbf{Type} & \textbf{Gloss} & {\small \texttt{\tag{adj}(\tag{comp})}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{subst}}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{advl}}} \\
				\midrule
				\textbf{A1} & \eng{good} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} \\
				\textbf{A2} & \eng{old} & {\qipa иске (искерәк)} & {\qipa иске (искерәк)} & {\qipa — (—)} \\
				\textbf{A3} & \eng{dead} & {\qipa үле (—)} & {\qipa үле (—)} & {\qipa — (—)} \\
				\textbf{A4} & \eng{basic} & {\qipa төп (—)} & {\qipa — (—)} & {\qipa — (—)} \\
				\bottomrule
			\end{tabular}
		}
	}

@@

			\vspace{-1em}
			\begin{center}
			\scalebox{1.05}{
				\begin{tabular}{lrrr}
				\toprule
					\textbf{Language} & \textbf{Forms} & \textbf{Precision} (\%) & \textbf{Recall} (\%) \\
				\midrule
					Kyrgyz & 500 &  &  \\
					Kazakh & 1000 & 98.61 &  57.98 \\
					Tatar & 1000 & 95.03 & 85.65 \\
					Kumyk & 500 & 96.57 & 69.11 \\
				\bottomrule
				\end{tabular}
			}
			\end{center}

@@

>>>>

	\headerbox{Development effort}{name=devel,column=4,row=0}{
	\scalebox{0.92}{
		\begin{tabular}{lllll}
			\toprule
			 & \textbf{Kyrgyz} & \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
			\midrule
			begun & Apr.\ 2011 & Dec.\ 2010 & Dec.\ 2011 & Oct.\ 2013 \\
			80\% cov.\ & Aug.?\ 2011 & Aug.\ 2012 & Aug.\ 2012 & Oct.\ 2013 \\\midrule
			time & 4 months & 19 months & 7 months & 1 week \\
			\bottomrule
		\end{tabular}
	}
		\begin{itemize}
			%\item Kyrgyz and Kazakh transducers begun in 2011
			%\item Kyrgyz transducer reached 80\% coverage summer of 2011
			\item (various periods of intermission, various rewrites)
			\item Kazakh transducer based on Kyrgyz transducer
			\item Kyrgyz transducer currently being rewritten based on insights gained while writing other Turkic transducers
			\item Kumyk transducer based on Kazakh, Tatar transducers:\\
			\tilda{}1 week to reach 80\% coverage, +1 week to reach 90\%
		\end{itemize}
		\vspace*{-2.5em}
		\hspace*{-2em}\includegraphics[width=0.9\textwidth,angle=270]{../kipchak-lrec-2013/graphs/kumyk-yoldash-cov}
		\vspace*{-2em}
	}

@@

		\headerbox{Categorisation}{name=adjectives,column=4,below=devel}{
		%\headerbox{Morphophonology}{name=morphophon,below=morphtrans}{
			\htwo{Adjectives}
			\begin{itemize}
				\item morphologically distinct adjective classes
				\item most sources claim:\ adjectives can be used substantively and adverbially
				\item Other Turkic transducers:\ \texttt{0}-derivation (\hilitetwo{overgenerates})
				\item but not all adjectives have all of the following:\\
				comparative forms, substantive readings, adverbial readings
				\item Our approach:\ categorisation 
				\item if properly categorised, only correct forms are analysed and generated
			\end{itemize}

		\scalebox{0.7}{
			\begin{tabular}{lllll}
				\toprule
				\textbf{Type} & \textbf{Gloss} & {\small \texttt{\tag{adj}(\tag{comp})}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{subst}}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{advl}}} \\
				\midrule
				\textbf{A1} & \eng{good} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} \\
				\textbf{A2} & \eng{old} & {\qipa иске (искерәк)} & {\qipa иске (искерәк)} & {\qipa — (—)} \\
				\textbf{A3} & \eng{dead} & {\qipa үле (—)} & {\qipa үле (—)} & {\qipa — (—)} \\
				\textbf{A4} & \eng{basic} & {\qipa төп (—)} & {\qipa — (—)} & {\qipa — (—)} \\
				\bottomrule
			\end{tabular}
		}

		\htwo{Adverbs}
			\begin{itemize}
				\item Certain adverbs have special attributive and ablative forms
				\item Mostly time adverbs
				%\item Mostly time adverbs, e.g.\\
				%{\qipa бүгүн} \eng{today}, {\qipa быйыл} \eng{this year}, {\qipa кечээ} \eng{yesterday}, {\qipa жана} \eng{just now}
				%\item Attributive forms:\\
				%{\qipa бүгүнкү} \eng{today's}, {\qipa быйылкы} \eng{this year's}, {\qipa кечээги} \eng{yesterday's}, {\qipa жанагы} \eng{just now}
				%\item Ablative forms:\\
				%{\qipa бүгүнтөн} \eng{from today}, {\qipa быйылтан} \eng{from this year}, {\qipa кечээтен} \eng{from yesterday}, {\qipa жанатан} \eng{from just now}
				\item Some also have noun readings:\ regular ablative, other cases:
				%{\qipa бүгүндөн} \eng{from today}, {\qipa быйылдан} \eng{from this year}
			\end{itemize}
			\scalebox{0.83}{
			\begin{tabular}{lllll}
				\toprule
				word & {\qipa бүгүн} & {\qipa быйыл} & {\qipa кечээ} &  {\qipa жана}\\
				gloss & \eng{today} & \eng{this year} & \eng{yesterday} & \eng{just now} \\
				\midrule
				\tag{attr} form & {\qipa бүгүнкү} & {\qipa быйылкы} & {\qipa кечээги} &  {\qipa жанагы}\\
				\tag{adv}\tag{abl} form & {\qipa бүгүнтөн} & {\qipa быйылтан} & {\qipa кечээтен} &  {\qipa жанатан}\\
				\tag{n}\tag{abl} form & {\qipa бүгүндөн} & {\qipa быйылдан} & — & — \\
				\bottomrule
			\end{tabular}
			}\vspace{0.2em}
				\codeex{
					\begin{spacing}{0.8}
					\texttt{LEXICON ADV-WITH-KI-ABL\\
\\
ADV ;\\
ADV-KI ;\\
ADV-ABL ;
					}
					\end{spacing}
					\vspace{0em}
				}
	}

@@

			\vspace{-1em}
			\begin{center}
			\scalebox{1.05}{
				\begin{tabular}{lrrr}
				\toprule
					\textbf{Language} & \textbf{Forms} & \textbf{Precision} (\%) & \textbf{Recall} (\%) \\
				\midrule
					Kyrgyz & 200 & 90.77 & 69.15 \\
					Kazakh & 1000 & 98.61 &  57.98 \\
					Tatar & 1000 & 95.03 & 85.65 \\
					Kumyk & 500 & 96.57 & 69.11 \\
				\bottomrule
				\end{tabular}
			}
			\end{center}

@@

		\headerbox{Other uses}{name=uses,column=1,below=morphophon}{
			\begin{itemize}
				\item HFST transducers are trivially converted to \textbf{spell checkers}
				\item Segmenter, e.g.\ {\qipa көргөзгөндөрдөнсүңбү} :\\
				 {\small \texttt{көр\hilitegrey{>}\{G\}\{A\}з\hilitegrey{>}\{G\}\{A\}н\hilitegrey{>}\{L\}\{A\}р\hilitegrey{>}\{D\}\{A\}н\hilitegrey{>}с\{I\}ң\hilitegrey{>}\{B\}\{I\}}}

				 %көр\hilitegrey{>}\{I\}ш\hilitegrey{>}\{G\}\{A\}н\hilitegrey{>}\{L\}\{A\}р\hilitegrey{>}\{I\}м\hilitegrey{>}\{D\}\{A\}н\hilitegrey{>}с\{I\}ң\hilitegrey{>}\{B\}\{I\}}}
			\end{itemize}
		}
